---
title: "PML_Project"
author: "JH"
date: "7/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PML_Project

# Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

# Goal of the project
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

# Data Analysis
First, load the data
```{r, cache= TRUE}
training <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"))
testing <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"))
```
There are a lot of parameters that contains most of NAs, get rid of the NAs parameters first.
```{r, cache= TRUE}
library(dplyr)
name_of_na <- names(which(apply(training, 2, function(x) sum(is.na(x)))==0))
trainingNoNA <- select(training,all_of(name_of_na))
testingNoNA <- select(testing,all_of(name_of_na[-length(name_of_na)]))
```
There are also a lot of parameters that only contain a few values, these parameters should also be getting rid of.
```{r, cache= TRUE}
name_of_null <- names(which(apply(training, 2, function(x) sum(x =="") == 0)))
training_NoNull <- select(trainingNoNA,all_of(name_of_null))
testing_NoNull <- select(testingNoNA,all_of(name_of_null[-length(name_of_null)]))
```
The classe classification should do nothing with the X and user_name, and since we are predicting every single movement, the time related parameter should be irrelevant and should be getting rid of.
```{r, cache= TRUE}
final_training<- training_NoNull[,-c(1:7)]
final_testing <- testing_NoNull[,-c(1:7)]
```
Then patition the final_training data set into the training and testing for model simulation process with the caret package.
```{r, cache= TRUE}
library(caret)
set.seed(123)
inTrain <- createDataPartition(final_training$classe,p = 0.75, list = FALSE)
thetraining <- final_training[inTrain,]
thetesting <- final_training[-inTrain,]
```
As the classes variable is the a classification variable, so we are going to use the decision trees method, k-Nearest neighbor and random forest methods to test out.

Before the classification, it is important to remove the highly correlated variables. Prepossessing with PCA will allow use to predict with less variable and less variance. So we will use pca preprocess for each method. 

Also for validation, I will use 10 fold cross validation with 3 repeats
# Decision Trees - rpart
Firstly, enable parallel running
```{r,results='hide'}
library(parallel)
library(doParallel)
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)
```

```{r,cache=TRUE}
set.seed(1231)
modeldt <- train(classe~., data = thetraining, method = "rpart",preProcess = "pca",trControl = trainControl(method = "repeatedcv", number = 10, repeats = 3,allowParallel = T))
confusionMatrix(as.factor(thetesting$classe),predict(modeldt,thetesting))$overall[1]
```
# K-nearest neighbors - knn
```{r, cache=TRUE}
set.seed(1232)
modelknn <- train(classe~., data = thetraining, method = "knn",preProcess = "pca",trControl = trainControl(method = "repeatedcv", number = 10, repeats = 3,allowParallel = T),tuneLength = 10)
confusionMatrix(as.factor(thetesting$classe),predict(modelknn,thetesting))$overall[1]

```


# Ramdon Forest - rf
```{r, cache=TRUE}
modelrf <- train(classe~., data = thetraining, method = "rf",preProcess = "pca",trControl = trainControl(method = "repeatedcv", number = 10, repeats = 3,allowParallel = T))
confusionMatrix(as.factor(thetesting$classe),predict(modelrf,thetesting))$overall[1]
```
# Conculsion and prediction
Comparing the three models, the knn and rf show high accuray and will be used for predicting the testing outcome.
```{r, cache=TRUE}
knnp <- predict(modelknn,final_testing)
rfp <- predict(modelrf,final_testing)
table(knnp,rfp)
```
We can that among the 20 elements, the disagreement of the two methods are 2 elements.



